The development of children's ability to track and predict turn structure in conversation

### 2015-12-17 paper/CasillasFrank-Anticip.tex
cuts in GD

Children begin to develop conversational turn-taking skills long before their first words \citep{bateson1975, hilbrink2015, jaffe2001, snow1977}. As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure. Until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors.

In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood and that spontaneous predictions are primarily driven by participants' expectation of an immediate response in the next turn. In making predictions about upcoming turn structure, children used a combination of lexical and prosodic cues; neither lexical nor prosodic cues alone were sufficient to support increased anticipatory gaze. We also found no early advantage for prosody over lexicosyntax, and instead found that children were unable to make above-chance anticipatory gazes in the \textit{prosody only} condition until age five. We discuss these findings with respect to the role of linguistic cues in predictions about upcoming turn structure, the importance of questions in spontaneous predictions about conversation, and children's developing competence as conversationalists.

@@

Importantly, we also found no evidence that lexical information alone is equivalent to full linguistic information for children, as has been shown before \citep{magyari2012, de-ruiter2006} and replicated in the current study for adult participants: In both experiments, children's performance was best in conditions when they had access to the full linguistic signal. Adults on the other hand, showed significant gains in anticipatory gaze switching in both conditions with lexical cues. If this effect arose in our data because children do not make predictions based on phonetically manipulated speech in conversational contexts, the current findings can be overturned by follow-up work controlling linguistic information through other means. But if not, there may be something specially informative about the combined prosodic and lexical cues to questionhood that boosts children's anticipations before they can use these cues separately. Even in adults, Torreira and colleagues (\citeyear{torreira2015}) were able to show that the trade-off in informativity between lexical and prosodic cues is more subtle in semi-natural (spliced) speech. The present findings are the first to show evidence of a similar effect developmentally.

\subsubsection{The question effect}

@@

%Because the form of a question can constrain the type of response that will come next (e.g., a location after a \textit{where} question), questions can even help listeners predict specific upcoming content in the next turn.

%Our results suggest that question turns start significantly affecting children's predictions between ages three and four, but further testing with finer-grained age samples with stimuli focused on specific communicative acts and linguistic cues is needed to better sketch out the developmental trajectory of this effect. For example, the effect size and age of emergence might differ by question type (e.g., \textit{wh-} vs. \textit{yes-no}) or the location of question-identifying cues within the unfolding utterance (early vs. late), but we would not be able to see it given the current design.

Prior work on children's acquisition of questions indicates that they may already have some understanding about question-answer sequences by the time they begin to speak: Questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years, even though the types and complexity of questions change throughout development \citep{casillas2016, fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation question frequency by individual and socioeconomic class \citep{hart1992}.} For the first few years, many of the questions directed to children are ``test' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?''), but this changes as children get older. Questions help caregivers to get their young children's attention and to ensure that information is in common ground, even if the responses are non-verbal or infelicitous \citep{bruner1985, fitneva2012, snow1977}. So, in addition to having a special interactive status (for adults and children alike), questions are a core characteristic of many caregiver-child interactions, motivating a general benefit for questions in turn structure anticipation.

Two important questions for future work are then: (a) How does children's ability to monitor for questions in conversation relate to their prior experience with questions? and (b) what is it about questions that makes children and adults more likely to anticipatorily switch their gaze to addressees? Other request formats, such as imperatives, compliments, and complaints make a response from the addressee highly likely in the next turn \citep{schegloff2007}. Rhetorical and tag questions, on the other hand, take a similar form to prototypical polar questions, but often do not require an answer. So, though it is clear that adults and children anticipated responses more often for questions than non-questions, we do not yet know whether their predictive action is limited to turns formatted as questions or is generally applicable to turn structures that project an immediate response from the addressee.

The question effect itself has implications for our current theories about turn prediction. While participants may always use linguistic information to predict upcoming speaker changes (as they have in the two studies presented here), they might not always use linguistic information to predict upcoming turn ends, as is assumed in metalinguistic measures of turn-end prediction (e.g., pressing a button while listening to speech: \citealp{torreira2015, magyari2012, de-ruiter2006}) and as has been the focus of other turn-end prediction studies (\citealp{ford1996, duncan1972}). Our results rather suggest that participants' \textit{spontaneous} predictions, at least while viewing third-party conversation, are the results of question-monitoring, presumably achieved by recruiting linguistic cues to questionhood from the unfolding signal. In other words, our results suggest that predictions are driven by what is \textit{beyond} the end of the current turn---that questions, not lexical cues, are sufficient for prediction at the first level of conversation monitoring.

There is at least one clear hypothesis that can bridge these apparently conflicting results: Listeners in spontaneous, first-person conversation may use multiple strategies in making predictions about upcoming turn structure, using more passive prediction to detect upcoming speaker transition (e.g., questions), and then switching into precise turn-end prediction mode (\`{a} la \citealp{de-ruiter2006}) when necessary. A flexible prediction system like this one allows listeners to continuously monitor ongoing conversation at a low cost while still managing to plan their responses and come in quickly when needed.

To test this hypothesis, which integrates findings from turn-structure prediction with multiple measures, age groups, and styles of linguistic control, it will be crucial to look at prediction from a first-person perspective. The results we present here are based on predictions about third-party conversation, which enables participants to follow interactions with no chance of actually participating. Although recent work has shown that similar anticipatory eye gazes do occur in spontaneous conversation \citep{holler2015}, more work is needed to determine if the same question advantage occurs, which linguistic cues seem to drive it, and whether participants switch into a ``precision'' mode when they detect imminent speaker change.

\subsubsection{Early competence for turn taking?}

@@

There are at least two major limitations to our work: Speech naturalness and participant role.Following prior work \citep{de-ruiter2006, keitel2013}, we used phonetically manipulated speech in Experiment 2, resulting in speech sounds that children don't usually hear in their natural environment. Many prior studies have used phonetically-altered speech with infants and young children \citep[cf.][]{jusczyk2000}, but almost none of them have done so in a conversational context. Future work could instead carefully script or cross-splice parts of turns to control for the presence or absence of linguistic cues for turn transition.

>>>>

Children begin to develop conversational turn-taking skills long before their first words \citep{bateson1975, hilbrink2015, jaffe2001, snow1977}. As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure. Until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors. In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood and that spontaneous predictions are primarily driven by participants' expectation of an immediate response in the next turn (e.g., for questions). In making predictions about upcoming turn structure, children used a combination of lexical and prosodic cues; neither lexical nor prosodic cues alone were sufficient to support increased anticipatory gaze. We also found no early advantage for prosody over lexicosyntax, and instead found that children were unable to make above-chance anticipatory gazes in the \textit{prosody only} condition until age five. We discuss these findings with respect to the role of linguistic cues in predictions about upcoming turn structure, the importance of questions in spontaneous predictions about conversation, and children's developing competence as conversationalists.

@@

Importantly, we also found no evidence that lexical information alone is equivalent to full linguistic information for children, as has been shown before \citep{magyari2012, de-ruiter2006} and replicated in the current study for adult participants. In both experiments, children's performance was best in conditions when they had access to the full linguistic signal. Adults on the other hand, showed significant gains in anticipatory gaze switching in both conditions with lexical cues. If this effect arose in our data because children do not make predictions based on phonetically manipulated speech in conversational contexts, the current findings can be overturned by follow-up work controlling linguistic information through other means. But if not, there may be something specially informative about the combined prosodic and lexical cues to questionhood that boosts children's anticipations before they can use these cues separately. Even in adults, Torreira and colleagues (\citeyear{torreira2015}) were able to show that the trade-off in informativity between lexical and prosodic cues is more subtle in semi-natural (spliced) speech. The present findings are the first to show evidence of a similar effect developmentally.

\subsection{The question effect}

@@

Prior work on children's acquisition of questions indicates that they may already have some understanding about question-answer sequences by the time they begin to speak: Questions make up approximately one third of the utterances children hear, before and after the onset of speech, and even into their preschool years, even though the types and complexity of questions change throughout development \citep{casillas2016, fitneva2012, henning2005, shatz1979}.\footnote{There is substantial variation question frequency by individual and socioeconomic class \citep{hart1992}.} For the first few years, many of the questions directed to children are ``test'' questions---questions that the caregiver already has the answer to (e.g., ``What does a cat say?''), but this changes as children get older. Questions help caregivers to get their young children's attention and to ensure that information is in common ground, even if the responses are non-verbal or infelicitous \citep{bruner1985, fitneva2012, snow1977}. So, in addition to having a special interactive status (for adults and children alike), questions are a core characteristic of many caregiver-child interactions, motivating a general benefit for questions in turn structure anticipation.

Two important questions for future work are then: (1) How does children's ability to monitor for questions in conversation relate to their prior experience with questions? and (2) what is it about questions that makes children and adults more likely to anticipatorily switch their gaze to addressees? Other request formats, such as imperatives, compliments, and complaints make a response from the addressee highly likely in the next turn \citep{schegloff2007}. Rhetorical and tag questions, on the other hand, take a similar form to prototypical polar questions, but often do not require an answer. So, though it is clear that adults and children anticipated responses more often for questions than non-questions, we do not yet know whether their predictive action is limited to turns formatted as questions or is generally applicable to turn structures that project an immediate response from the addressee.

% The question effect itself has implications for our current theories about turn prediction. While participants may always use linguistic information to predict upcoming speaker changes (as they have in the two studies presented here), they might not always use linguistic information to predict upcoming turn ends, as is assumed in metalinguistic measures of turn-end prediction (e.g., pressing a button while listening to speech: \citealp{torreira2015, magyari2012, de-ruiter2006}) and as has been the focus of other turn-end prediction studies (\citealp{ford1996, duncan1972}). Our results rather suggest that participants' \textit{spontaneous} predictions, at least while viewing third-party conversation, are the results of question-monitoring, presumably achieved by recruiting linguistic cues to questionhood from the unfolding signal. In other words, our results suggest that predictions are driven by what is \textit{beyond} the end of the current turn---that questions, not lexical cues, are sufficient for prediction at the first level of conversation monitoring.

% There is at least one clear hypothesis that can bridge these apparently conflicting results: Listeners in spontaneous, first-person conversation may use multiple strategies in making predictions about upcoming turn structure, using more passive prediction to detect upcoming speaker transition (e.g., questions), and then switching into precise turn-end prediction mode (\`{a} la \citealp{de-ruiter2006}) when necessary. A flexible prediction system like this one allows listeners to continuously monitor ongoing conversation at a low cost while still managing to plan their responses and come in quickly when needed.

% To test this hypothesis, which integrates findings from turn-structure prediction with multiple measures, age groups, and styles of linguistic control, it will be crucial to look at prediction from a first-person perspective. The results we present here are based on predictions about third-party conversation, which enables participants to follow interactions with no chance of actually participating. Although recent work has shown that similar anticipatory eye gazes do occur in spontaneous conversation \citep{holler2015}, more work is needed to determine if the same question advantage occurs, which linguistic cues seem to drive it, and whether participants switch into a ``precision'' mode when they detect imminent speaker change.

\subsection{Early competence for turn taking?}

@@

There are at least two major limitations to our work: Speech naturalness and participant role. Following prior work \citep{de-ruiter2006, keitel2013}, we used phonetically manipulated speech in Experiment 2, resulting in speech sounds that children don't usually hear in their natural environment. Many prior studies have used phonetically-altered speech with infants and young children \citep[cf.][]{jusczyk2000}, but almost none of them have done so in a conversational context. Future work could instead carefully script or cross-splice parts of turns to control for the presence or absence of linguistic cues for turn transition.
